### 2013-03-27 tex/ms.tex
intro paragraphs

\documentclass[12pt,preprint]{aastex}

% Import the natbib package for bibliography inclusion. To get it to work right, first typeset the bibliography (BibTex), Command-Shift-B.
% Then, typeset the document with LaTeX, Command-Shift-L or just Command-T.
% To get proper reference citations using aastex (preprint manuscripts), first typeset as emulateapj with BibTex to create the BibTex
% files and then typeset again as aastex to get the manuscript with references.
\usepackage{natbib}
\usepackage{epsfig,graphicx,color}
\usepackage{longtable}
\usepackage{amsmath}

\bibliographystyle{apj}
% A few colors to replace the defaults for certain link types
\definecolor{orange}{cmyk}{0,0.4,0.8,0.2}
\definecolor{darkorange}{rgb}{.71,0.21,0.01}
\definecolor{darkgreen}{rgb}{.12,.54,.11}

%-----------------------------------------------------------------------------
% The hyperref package gives us a pdf with properly built
% internal navigation ('pdf bookmarks' for the table of contents,
% internal cross-reference links, web links for URLs, etc.)
\usepackage{hyperref}
\hypersetup{pdftex,  % needed for pdflatex
  breaklinks=true,  % so long urls are correctly broken across lines
  colorlinks=true,
  urlcolor=blue,
  linkcolor=darkorange,
  citecolor=darkgreen,
  }

%Define this shorthand for scientific notation.
\newcommand{\expnt}[2]{\ensuremath{#1 \times 10^{#2}}}   % scientific notation
\newcommand{\x}{{\bf x}}
\newcommand{\X}{{\bf X}}
\newcommand{\y}{{\bf y}}
\newcommand{\z}{{\bf z}}
\newcommand{\new}{red}
\newcommand{\ptr}{p_{\rm tr}}
\newcommand{\pte}{p_{\rm te}}
\newcommand{\Ex}{\mathbf{E}}
\newcommand{\Prob}{\mathbf{P}}
\newcommand{\PrfL}{\widehat{P}_{\rm{RF,}\mathcal{L}}(y|\x)}
\newcommand{\PrfLx}{\widehat{P}_{\rm{RF,}\mathcal{L}\cup\x'}(y|\x)}
\newcommand{\TreeL}{\theta_{b, \mathcal{L}}(y|\x)}
\newcommand{\TreeLx}{\theta_{b, \mathcal{L}\cup\x'}(y|\x)}
 \def\hipp {{\it Hipparcos~}}

\newcommand{\dd}{\mathrm{d}}

\newcommand{\fobs}{f_i}
\newcommand{\sobs}{s^2_i}

\newcommand{\ftrue}{f_i^*}

% Commands for annotating the docs with fixme and inter-author notes.  See
% below for how to disable these.
%
% Define a \fixme command to mark visually things needing fixing in the draft,
% as well as similar commands for each author to leave initialed special
% comments in the document.
% For final printing or to simply disable these bright warnings, symlink
% (there's a target 'dist_on' in the makefile that does this) the file
% macros_state.tex to macros_off.tex

\newcommand{\fixme}[1] { \textcolor{red} {
{\fbox{ {\bf FIX}
\ensuremath{\blacktriangleright \blacktriangleright \blacktriangleright}}
{\bf #1}
\fbox{\ensuremath{\blacktriangleleft \blacktriangleleft \blacktriangleleft}
} } } }

% And similarly, one (less jarring, with fewer symbols and no boldface) command
% for each one of us to leave comments in the main text.
\newcommand{\james}[1] { \textcolor{blue} {
\ensuremath{\blacklozenge} {\bf james says:}  {#1}
\ensuremath{\blacklozenge} } }

\newcommand{\hogg}[1] { \textcolor{darkorange} {
\ensuremath{\blacksquare} {\bf hogg says:}  {#1}
\ensuremath{\blacksquare} } }

\newcommand{\dan}[1] { \textcolor{darkgreen} {
\ensuremath{\bigstar} {\bf dan says:}  {#1}
\ensuremath{\bigstar} } }

\newcommand{\joey}[1] { \textcolor{red} {
\ensuremath{\clubsuit} {\bf joey says:}  {#1}
\ensuremath{\clubsuit} } }%\documentclass[manuscript]{aastex}

\begin{document}
%%%%%%%% TITLE, AUTHORS, PUBLICATION STATUS, and ABSTRACT %%%%%%%%%

\shorttitle{Fitting the Undetected}
\shortauthors{Richards, Hogg, Long, Foreman-Mackey}
\title{Fitting the Undetected:  Light-curve Feature Estimation in the Presence of Non-detections}
\author{
Joseph W. Richards\altaffilmark{1, 2, 3, *},
David W. Hogg\altaffilmark{4, 5},
James P. Long\altaffilmark{2},
Daniel Foreman-Mackey\altaffilmark{4}
}
%
\altaffiltext{1}{Astronomy Department, University of California, Berkeley, CA, 94720-7450, USA}
\altaffiltext{2}{Statistics Department, University of California, Berkeley, CA, 94720-7450, USA}
\altaffiltext{3}{wise.io, 2150 Shattuck Ave., suite 525, Berkeley, CA, 94704, USA}
\altaffiltext{4}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA}
\altaffiltext{5}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
\altaffiltext{*}{\texttt{jwrichar@stat.berkeley.edu}}

% Note the status (in preparation, MNRAS submitted date, ApJ accepted date, in press etc).
\slugcomment{DRAFT 2013-03-26 / in preparation}

\begin{abstract}
In multi-epoch imaging surveys, faint variable sources will not be
detectable at all epochs.  In principle, the best approach is to
perform photometric measurements (forced photometry) anyway.
%
In practice, most surveys only provide detection upper limits at some
epochs; in some not even upper limits but just the ``null''
information that the star was not detected.
%
Here we demonstrate with real data on periodic variables from the
All-Sky Automated Survey (ASAS) that these null observations can
in fact be very useful in fitting and parameter estimation.
%
We build a novel light-curve model that includes the per-epoch
detection thresholds as a set of latent parameters, one per data
point; the best-fit model is the maximum-likelihood model after
marginalizing out the latent thresholds.
%
The key assumption is that the detection thresholds are drawn from a
stationary distribution (the parameters of which are free to vary).
%
We demonstrate that this censored-data model outperforms traditional
fitting---in which null observations are discarded from the
fits---in estimating both periods and amplitudes for stars subject to
heavy censoring.
%
In particular, our model eradicates a failure mode in which
traditional fitting obtains an integer fraction of the correct period.
\end{abstract}

%\keywords{stars: variables: general -- methods: data analysis -- methods: statistical -- techniques: photometric}

\section{Introduction}
\label{sec:intro}

Telescopes don't make catalogs, they make images (Hogg \& Lang 20xx).
Nonetheless, some of the most important astronomical projects have produced enormously productive catalogs
(the \textsl{SDSS}, for examle CITATION).
There is a growing understanding that precise measurements of astronomical sources have to be made directly in the imaging data
(CITE Lang, Anderson, something in weak lensing):
It is impossible to pass the full information content of the imaging to the end user via a catalog
(which by necessity is built with point estimates of the quantities of interest).

The measurements of importance for variable-star science are relatively simple.
Unlike the measurements used for astrometry, extragalactic transients, and weak lensing,
where two-dimensional (morphological) information matters,
when variable stars are being observed, all the end user wants is accurate photometry of each source at each epoch.
One could hope that catalog-level information is sufficient:
the detailed image morphology is not required for lightcurve extraction, except when fields are crowded.
However, there are still limitations to working at catalog level:
Catalog entries (nowadays) usually come with noise variance estimates,
but they don't usually give detailed information about the locations or influence of
bad detector pixels, cosmic rays, and flat-field variations;
all these things might be needed to construct a precise photometric noise model or to understand residuals or artifacts.
More importantly, catalogs are usually subject to hard \emph{censoring};
the catalogs (nowadays) only include measurements that are deemed (by the software or catalog makers)
to be \emph{signficant detections}.

This could change.
In principle a catalog could contain a measurement at \emph{every} epoch for any source that is detected at \emph{any} epoch.
This kind of ``forced photometry'' would include many non-significant measurements,
but it would come closer to passing forward all the information in the imaging to users working at catalog level.
We recommend this approach to the catalog makers of the future (especially those making the \textsl{LSST} catalogs).

Less good than the comprehensive forced-photometry approach is to report, when measurements are not significant,
a well-defined ``upper limit'' for the brightness of each source.
This is rarely done, and upper limits are hard to use, but upper limits are an improvement over standard practice.
Standard practice is to \emph{report nothing at all} when a source is not detected significantly.
Indeed, it is often hard, when presented with a multi-epoch catalog,
to determine whether the reason a particular source is missing from the imaging at a particular epoch
is because the source was faint or because the source fell outside the imaging field of view!

In this contribution we look at improving catalog-level time-domain astronomy in the presence of \emph{censored data}.
The problem we consider is that of period and amplitude modeling for large-amplitude variable stars in a data set
in which \emph{no information is reported at all} at the epochs at which the star is not significantly detected.
Furthermore, we will imagine that we also have \emph{no information at all} about the sensitivity of the telescope
at the time of each non-detection; that is, we have no upper limits and we can't reconstruct them from any extant meta-data.
The non-detections are literally database ``null'' values.
This set-up may seem artificial, but it is in fact the situation we find ourselves in for various surveys,
including the ASAS survey, data from which we will model below.
What we will show is that \emph{the ``null'' non-detections are in fact very informative};
amplitude and period fitting will be wrong in general if the non-detections are ignored.

The benefit of using the non-detections is large, but the cost is also large:
Use of these ``null'' data involves non-trivial modeling,
and large numbers of latent parameters to be inferred and marginalized out.

To make this all concrete, consider the photometric data from the \textsl{ASAS} survey
depicted in \figurename~\ref{fig:example} for one Mira variable.
The photometric detections are shown and so are the epochs at which there are nulls (or non-detections).
The standard

...as an example, consider \figurename~\ref{fig:example}...

\joey{What people actually do, what problems that could cause, and what people should do}

A photometric survey (think of it as operating in a single photometric
bandpass for now) scans over the celestial position of a particular
variable star  a large number of times $t_i$.  At each of these
times, the imaging data are analyzed by software, which
treats each scan as a \emph{completely independent} survey.  That is,
when analyzing the data from time $t_i$, none of the information from
any of the other times is used in any way.

At each time $t_i$, the star is either detected by the
software or not.  When it is detected, the software returns
a (possibly bad) flux value $f_i$ and (likely bad) uncertainty
variance $s_i^2$.  When the source is \emph{not} detected,
nothing is reported---in this case, because the software is not
awesome, it doesn't see any reason to report anything at all, since on
the non-detect passes, it has no inkling that this patch of the sky is
interesting in any way.  Welcome to the desert of the real.

Beyond this, there are two additional issues.  The first is that each
night of imaging is different in an unreported and unknown way.  That
is, there is different transparency, sky brightness, and point-spread
function.  So the detection limit, or completeness level, or censoring
of the catalog is different every night.  The second issue is that
either because we are getting the data from a non-generous source, or
because the conditions change rapidly and unpredictably, we can't
analyze all the sources in a finite patch simultaneously.  For each
variable star of interest, we \emph{only} get the data on that star
itself.

\joey{RRL end at 120 kpc in Sesar et al. paper}

\joey{discuss in context of: parameter estimation (amplitude / period), discovery (we're gonna find a lot more variable sources if we take the non-detections into account), and classification; lossiness}

\joey{let's make a point to stress variable discovery}

\joey{Numbers of LCs affected: Stripe 82, ASAS}

\section{The Model}
\label{sec:model}

As a reminder, we are modeling light curve data in the presence of non-detections, which are epochs of observation in which no detection of the source of interest was made.

\subsection{Preliminaries}
\label{ss:prelim}

For each astronomical object, a photometric survey measures a multi-epoch light curve over $N$ (typically unevenly spaced) epochs.  At each epoch $i$, with associated time $t_i$, the survey takes an exposure at the location of the object and either (a) detects the object and records an estimate of its photon flux, $\fobs$ and the  variance of the statistical uncertainty in that estimate, $\sobs$, or (b) fails to detect the object.  In the latter case, most modern surveys either record a reference value to signify that no detection was made (as is done in ASAS) or an estimate of the upper detection limit, $b_i$, which is the minimum brightness that could have been detected.

There are many reasons that a source might not show up in a catalog.  These include:
\begin{itemize}
\item low S/N of the source, due either to a higher noise level or a fainter signal,
\item the source falling outside of the detection window (e.g., near chip edge),
\item occulting of the source by an artifact of the detector (e.g., hot pixels, masked out), or
\item the source was out-shone by an intervening object (e.g., asteroid, comet, variable star, airplane, etc.).
\end{itemize}
Here, we present a statistical model that can be used to detect variable sources and model their variability using multi-epoch light curves containing epochs of non-detection.  Previous efforts to detect and model variability using multi-epoch photometry have typically ignored non-detections (REFs) or used them in an ad-hoc manner lacking statistical rigor (REFs).  An exception is \citet{2009AJ....137.4400L}, who measure proper motions of sources in SDSS falling below the detection limit.

In this paper, we will assume that $\fobs$ and $\sobs$ take on real-valued numbers in epochs for which the source was detected and receive the reference value {\tt NA} in epochs where no detection was made. For notational convenience we assemble all the light curve data for one source into a data set $D$ given by
\begin{eqnarray}\displaystyle
D &\equiv& \{D_i\}_{i=1}^N
\\
D_i &\equiv& (\fobs, \sobs)
\quad ,
\end{eqnarray}
where $(\fobs, \sobs)$ are the light curve measurements  at $t_i$.  The goal, for each astronomical object, is to construct the likelihood for the data $D$, given a set of model parameters that describe the variability of the object and the characteristics of the observations, and then to maximize the likelihood with respect to the model parameters.% or to use it in further inference.

\subsection{Modeling Non-Detections}
To model the mean brightness of the light curve as a function of time, we use a multiple-harmonic Fourier model with angular oscillation frequency $\omega$,
\begin{eqnarray}\displaystyle
\mu_i' &=& A_0 + \sum_{k=1}^K A_k\, \sin (t_i \, \omega  k) + B_k\, \cos (t_i \, \omega  k)
\quad ,\label{eq:fourier}
\end{eqnarray}
where $\mu_i'$ is the model brightness estimate in magnitude at time $t_i$, $\sqrt{A_k^2 + B_k^2}$ is the amplitude of the $k^{\rm th}$ harmonic of the frequency $\omega$ and $\tan^{-1}(\frac{B_k}{A_k})$ is the relative phase offset of harmonic $k$.  The number of harmonics, $K$, can either be fixed or treated as a free parameter over which to optimize the likelihood.

We model the brightness in magnitude because periodic variation in magnitude is more sinusoidal than in flux. However since errors more approximately Gaussian in flux space, we convert the $\mu_i'$ to flux  using
\begin{equation*}
\mu_i = 10^6 \times 10^{-0.4 \mu_i'}
\end{equation*}
Now both the model predictions $\mu_i$ and the data $(f_i,s_i)$ are in flux.

In addition to the expected brightness in Equation \ref{eq:fourier}, we assume that there is uncertainty or inappropriateness in the model, leading to a \emph{fractional model variance} $\eta^2$ at each point (assumed constant but easily generalized) as well as random variation due to photon arrival counts, assumed normal mean 0 with variance $\sigma_i^2$. The reported uncertainties $s_i^2$ are an estimate of the $\sigma_i^2$, see Subsection \ref{sec:uncertainty}. This leads to a true flux measurement of
\begin{eqnarray}\displaystyle
\ftrue = \mu_i + e_i
\label{eq:truefluxdef}
\end{eqnarray}
where $e_i \sim N(0,\sigma_i^2 + \eta^2\,\mu_i^2)$. Recall that $f_i^*$ is a latent variable because we will not see it when an observation is censored. We instantiate a latent variable, $b_i$, to represent the detection limit, in units of flux, at epoch $i$.  The $b_i$ parameter is essential because it probabilistically constrains the possible values of the mean brightness, $\mu_i$, of the light curve when there is a non-detection.  By employing a hierarchical model for the distribution of $b_i$, we can fully utilize all of the information encoded in both the detections and non-detections when computing the data likelihood.   At each epoch, we connect the observed flux, $\fobs$, to the latent variable $\ftrue$ by
\begin{eqnarray}\displaystyle
\fobs &=& \left\{\begin{array}{ll}
  \ftrue & \mbox{if $\ftrue \ge b_i$} \\
  \texttt{NA} & \mbox{if $\ftrue < b_i$} \\
\end{array} \right.
\end{eqnarray}
The observed flux is \texttt{NA} only when the true flux, $\ftrue$, is below the detection limit $b_i$.

Using the model above, the likelihood of the observed $\fobs$ given $\sigma^2_i$, $\theta$ and $I$ (after integrating out the unknown $b_i$) is
\begin{eqnarray}\displaystyle
p(\fobs |\sigma^2_i,\theta,I) &=& \left\{\begin{array}{ll}
  N(\fobs | \mu_i,  \sigma^2_i + \eta^2\,\mu_i^2)\,  \int_{-\infty}^{\fobs} p(b_i | \theta)\, \dd b_i & \mbox{if $\fobs \ne \texttt{NA}$} \\
  \int_{-\infty}^{\infty} \int_{-\infty}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + \eta^2\,\mu_i^2)\, p(b_i | \theta)\, \dd \ftrue\, \dd b_i & \mbox{if $\fobs = \texttt{NA}$} \\
\end{array}\right.\label{eq:mlik}
\\
p(b_i|\theta) &=& N(b_i|B,V_B)
\label{eq:bprior}
\\
\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, \eta^2, B, V_B, \cdots) \quad ,
\end{eqnarray}
where we have introduced the hyperparameters $B$ and $V_B$ for the Gaussian prior distribution of $b_i$.  In Equation \ref{eq:mlik}, in the epochs for which a detection was made ($\fobs \ne \texttt{NA}$), we marginalize over the unknown $b_i$ up to the observed flux $\fobs$, enforcing that the detection limit be fainter than the observed brightness. Computationally, this requires a call to a Gaussian probability density function (pdf), a call to a Gaussian cumulative distribution function (cdf), and a multiplication. In the epochs for which no detection was made ($\fobs = \texttt{NA}$), we compute the probability that the unknown detection threshold $b_i$ is greater than the unknown true flux $f_i^*$. When both $b_i$ and $f_i^*$ are normally distributed, this can be computed using a single call to a Gaussian cdf.

In the above, we have assumed no extra information on each of the $b_i$ values besides the knowledge of whether a detection was made at that epoch.  Hence, we draw, in Equation \ref{eq:bprior}, each $b_i$ value from a global prior distribution which is the same at all epochs.  If instead, we are given an estimate of $b_i$ plus its error distribution for each epoch (which, in principle can be inferred from the raw telescope images), we can replace Equation \ref{eq:bprior} with a different distribution per epoch.  In the case that the $b_i$ are assumed to be completely known (without error), the data likelihood of $\fobs$ becomes
\begin{eqnarray}\displaystyle
p(\fobs |\sigma^2_i,\theta,I, b_i) &=& \left\{\begin{array}{ll}
  N(\fobs | \mu_i,  \sigma^2_i + \eta^2\,\mu_i^2)\,  I(\fobs \ge b_i) & \mbox{if $\fobs \ne \texttt{NA}$} \\
 \int_{-\infty}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + \eta^2\,\mu_i^2)\, \dd \ftrue & \mbox{if $\fobs = \texttt{NA}$} \\
\end{array}\right.\label{eq:mlik_s}
\quad ,
\end{eqnarray}
where the boolean indictor function, $I(\fobs \ge b_i)$, is 1 if $\fobs \ge b_i$ and 0 otherwise.

\subsection{Modeling Uncertainty in Uncertainties}
\label{sec:uncertainty}
Instead of assuming that $\sobs$ is a perfect, error-free measurement, we can probabilistically connect it to the true photon flux uncertainty variance $\sigma^2_i$ through a Gamma likelihood.  Our model probability of observed $\sobs$ given $\fobs$, $\sigma^2_i$, $\theta$ and $I$, is
\begin{eqnarray}\displaystyle
p(\sobs | \fobs, \sigma^2_i, \theta,I) &=& \left\{\begin{array}{ll}
  \Gamma (\sobs | \sigma^2_i, V_{\sigma} ) & \mbox{if $\fobs \ne \texttt{NA}$} \\
 I(\sobs = \texttt{NA})& \mbox{if $\fobs = \texttt{NA}$}
\end{array}\right.\label{eq:slik}
\quad ,
\end{eqnarray}
where $\Gamma(x | m, V)$ is the standard Gamma distribution parameterized by its mean $m$ and variance $V$\footnote{The $\Gamma(x | m, V)$ pdf is $ p(x ) = (\theta^k \Gamma(k))^{-1}  x^{k-1} e^{x/ \theta}$, for $x \ge 0$, where $\theta = V / m$ and $k  = m^2 / V$.}.
The  model hyperparameter $V_\sigma$ represents the
variance in the distribution of reported uncertainties around the true
uncertainty.  Smaller values of $V_\sigma$ represent higher credence given to the reported uncertainties whereas larger values mean that our model is given more flexibility to use observation uncertainties that are far from the reported values.  Ultimately, the model will choose the value of $V_\sigma$, for each light curve, that maximizes the data likelihood.

To complete our model of uncertainties on the reported observational error variances, $s_i^2$, we must place a prior distribution on the true uncertainties, $\sigma_i^2$.  We choose to place a Gamma prior on the true photon flux uncertainties, 
\begin{eqnarray}\displaystyle
p(\sigma^2_i|\theta, I) &=& \Gamma(\sigma^2_i | S, V_S)
\label{eq:sigma_prior}
\end{eqnarray}
adding two parameters, $S$ and $V_S$, to our model.
\james{should probably justify choice of some priors here / include some plots and/or refs. gamma priors and normal priors.}  This puts us in a position to write down our full likelihood model for light curves containing censored observations.

\subsection{The Full Model}
\label{sec:full}
Putting it all together, we can write down the likelihood of the data, $D_i$, for the parameters $\theta$, on a single epoch, as
\begin{eqnarray}\displaystyle
\label{eqn:lik}
p(D_i|\theta,I) &=& \int_0^{\infty} p(\fobs |\sigma^2_i,\theta,I)\, p(\sobs | \fobs, \sigma^2_i,\theta,I)\, p(\sigma^2_i | \theta, I)\,\dd \sigma^2_i
\\
\label{eqn:likpar}
\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, \eta^2, B, V_B, V_{\sigma}, S, V_S) \quad ,
\end{eqnarray}
where we have inserted the expressions from Equations (\ref{eq:mlik}), (\ref{eq:slik}), and (\ref{eq:sigma_prior}), and integrated over the nuisance parameter $\sigma^2_i$.

Finally if we assume that the data collected at each epoch are independent given the model parameters, we have that
\begin{eqnarray}\displaystyle
\label{eqn:likfull}
p(D|\theta,I) &=& \prod_i p(D_i|\theta,I)
\quad.
\end{eqnarray}
This is the likelihood for the entire
data set (all the measurements and non-detections of this star from
all the epochs, as delivered by the survey) given the $2K + 9$ parameter vector $\theta$.
Figure \ref{fig:graph} displays a graphical model highlighting dependencies and 
independence assumptions. This model
``correctly'' or at least ``justifiably'' uses all of the information
available, without making strong assumptions about the survey or its
veracity.

\subsection{Optimization in Practice}
\label{ss:optim}

Using the light curve data, $D$, from a particular source, our goal is to infer the parameters that describe the variability of that object.  We do so by maximizing the likelihood in Equations (\ref{eqn:lik}), and (\ref{eqn:likfull}) with respect to the parameter vector in Equation (\ref{eqn:likpar}).  The maximum likelihood estimate (MLE), which we call $\widehat{\theta}$ is the best estimate of the vector of model parameters that explain the variability of the object, given all of the data (observed and censored) and the assumed model.  The MLE allows us to then infer parameters of physical interest, such as the period and amplitude of oscillation.

To maximize the data likelihood with respect to $\theta$, we use the {\tt fmin} routine in the {\tt Python} module {\tt scipy.optimze}.  The {\tt fmin} routine uses a a Nelder-Mead simplex algorithm to optimize a function with respect to its arguments starting from some initial guess, $\theta_0$.  The convergence of the algorithm is highly sensitive to the value of $\theta_0$.  To find the MLE parameters for each light curve, we use {\tt fmin} to minimize the negative log likelihood (negative natural log of Equation \ref{eqn:likfull}).  We use a standard Lomb-Scargle Fourier analysis (ignoring any epochs of non-detection) to initialize the search algorithm, deriving an initial guess for $\omega$ as the frequency corresponding to the maximal Lomb-Scargle spectral density, and initializing $A_0$, and $\{A_k, B_k\}_{k=1}^K$ using standard least squares.  In practice, we run the {\tt fmin} routine multiple times using different initial values of the angular frequency corresponding to the frequencies of the most significant spectral density peaks and values of one-half those frequencies.  Initial values for the other parameters are set at $\eta^2 = 0.2$, $B = V_B = 2\cdot\min(\fobs)$, and $ V_{\sigma} = S = V_S = \textrm{med}(\sobs)$.

Replacing likelihood with fast version where we trust $s_i^2$ for observed data.

\section{Experiments}
\label{sec:experiments}

\joey{Note: ASAS provides periods and amplitude for all of the objects in ACVS!  We can and should use this as a comparison set}

\joey{From a brief reading of a few Mira papers, it seems that some papers throw away Miras for which no trough of the LC was observed.  This seems silly.  I have an idea of the P-A plots from Hipparcos.  Should be a good comparison set.  Also, we can consider applying this to Hipparcos, which has publicly available data.}

\subsection{Simulating Faint Mira Variables}
\label{ss:mirasim}

In this first experiment, we begin with a well-observed Mira variable from the ASAS Catalog of Variable Stars (ACVS, \citealt{acvs}) , ASAS 235627-4947.2.  This star has a pulsation period of 266.6286 days and a Lomb-Scargle amplitude of 2.38 mag (V band).  Note there are no non-detections in the ASAS light curve.

To simulate faint Mira stars from ASAS 235627-4947.2, we use the following procedure:
\begin{enumerate}
\item Convert the observed magnitudes $m_i$ and uncertainty variances $s^2_{m,i}$ to fluxes $f_i$ and flux uncertainty variances $s^2_{f,i}$ (we assume a V-band zero-point of $3.67\times10^{-9}$ erg/s/cm$^2$/\AA)
\item For a given flux dimming parameter, $d$, sample the new fluxes, $\tilde{f}_i$, from a Gaussian distribution centered around $f_i / d$ with standard deviation sampled from the empirical ASAS distribution of $s^2_{f,i} | f_i$. \joey{I found that relationship to be linear with a slope $\approx 1$, which is probably BS since the ASAS mag errors are all crap.  So I added a reference value to each flux error to ensure a mag limit $\approx$ 14.5 mag.}
\item Fluxes that are not at least 5$\sigma$ above zero are denoted as non-detections and their flux estimates (and errors) are censored.
\end{enumerate}
Folded light curves of ASAS 235627-4947.2, dimmed by four different values of $d$, are plotted in Figure \ref{fig:mirasim}.

\subsection{Results of Fitting Miras}

\section{Results: ASAS Light Curves}
\label{sec:results}

We use the methodology of \cite{2011arXiv1106.2832R} to select the top ASAS Mira  and RR Lyrae, Fundamental Mode candidates.  Using a posterior probability threshold of 0.8 gives us 1720 Mira  and 1029 RR Lyrae candidates.

 \subsection{Mira Variables}

Figure \ref{fig:miraP} shows the estimated periods for standard Lomb-Scargle against the periods from our censored Fourier model for 2538 likely Mira variables chosen from the MACC catalog (these were chosen such that $P$(Mira) $>0.5$.)  For 12.6\% of these objects, the censored period is twice that of the Lomb-Scargle period, while for 0.51\% it is triple (show examples).  This occurs because of XYZ.  For 85.5\% of all objects, the periods are consistent to $\pm$5\%.

Figure \ref{fig:miraPA} shows the period-amplitude relationship for the set of Mira variables, with periods and amplitudes estimated by standard Lomb-Scargle (left) and the censored model (right).  Results show that XYZ.

 \subsection{RR Lyrae Variables}

  Show P - A relationship before and after using the method

There is an RRL P - A relationship (strong linear anti-correlation)

As a selection criterion (simulated data?)

\section{Discussion}
\label{sec:discussion}

What people can do, starting from the data-taking procedure.

limitations

\acknowledgements It is a pleasure to thank
%
  Josh Bloom (Berkeley) and
  Nat Butler (ASU)
%
for valuable discussions and comments.
%
Partial support for this project was provided by \hogg{[JWR: PUT GRANT
  NUMBERS HERE]}, NASA grant NNX12AI50G, and NSF grant IIS-1124794.

\bibliography{non_detect}

\clearpage
\begin{figure}
  \begin{center}
\includegraphics[angle=0,width=6.5in]{../plots/fig_045832-0604_1.png}
  \end{center}
  \caption{A comparison of the best fit obtained by the traditional method (left) and our censored data model (right) on the Mira variable ASAS 045832-0604.1.  The top row shows the respective model fits to the raw light curves while the bottom row displays the period-folded data and models.  The black dots denote the observed data and the red vertical dashes mark the epochs of null observation.  The traditional fitting method clearly obtains half the true period while our method recovers the correct period.  The traditional method fails because it identifies a reasonable high-frequency fit to the observed data which completely ignores the null observations.  Note that alternate cycles of the traditional model fit contain no observed data.  By contrast, our method fully accounts for the null observations using a probability model.\label{fig:badfit}}
\end{figure}

\clearpage
\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth,%
      trim=3cm 12cm 5cm 3cm, clip=true]{gm/gm.pdf}
  \end{center}
  \caption{The graphical representation of the probability distribution.\label{fig:graph}}
\end{figure}

\clearpage
\begin{figure}
\begin{center}
\includegraphics[angle=0,width=6.5in]{../plots/mira_simulated.pdf}
\end{center}
\caption{ Folded light curves for data simulated from the Mira variable ASAS 235627-4947.2 using the prescription in \S\ref{ss:mirasim}.  Blue tick marks along the bottom axis denote phases where there were non-detections.  As the original flux measurements are dimmed by higher factors, the troughs of the sinusoidal light curve are censored, resulting in an incomplete view of the data.  \hogg{This figure needs to show the OLD and NEW best-fit lightcurve models too!}\label{fig:mirasim}}
\end{figure}

\clearpage
\begin{figure}
  \begin{center}
    [\hogg{PUT FIGURE HERE}]
  \end{center}
  \caption{A real-data example.  This is the same source as that shown in \figurename~\ref{fig:badfit} but now showing the results of...\label{fig:goodfit}}
\end{figure}

\clearpage
\begin{figure}
  \begin{center}
    [\hogg{PUT FIGURE HERE}]
  \end{center}
  \caption{Another few real-data examples; in each case the method presented here either matches the traditional method or is substantially better.\label{fig:examples}}
\end{figure}

\clearpage
\begin{figure}
\begin{center}
\includegraphics[angle=0,width=6.5in]{../plots/mira_periods_LS_vs_Censored.pdf}
\end{center}
\caption{ .  \label{fig:miraP}}
\end{figure}

\clearpage
\begin{figure}
\begin{center}
\includegraphics[angle=0,width=6.5in]{../plots/mira_period_amplitude_relation.pdf}
\end{center}
\caption{ .  \label{fig:miraPA}}
\end{figure}

\end{document}